---
title: "Annotating mineralization and environment"
output: html_notebook
---

Loading the all.fams.csv, combine it with the biomineralization data.  And also annotate it with the info of whether they are planktonic or benthic taxa.
The workflow would be to create unique 

```{r load packages and database}

library(tidyverse)
library(googlesheets)
library(taxize)
library(rentrez)
library(here)
source(here("Scripts","sequential.join.r"))

# set_entrez_key("")
# Sys.getenv("ENTREZ_KEY")

all.fams <- read_csv(here("Input","all.fams.tax.csv"))

all.hashes.found <- read_csv(here("Input","all.taxonomy.20190130.csv"))

biom <- read_csv(here("Input","Biomineralization.csv"))

```

So the workflow is going to be to try and get the information from each taxonomical rank. So I will create a function that takes 2 datasets, and binds them sequentially. For each join, it will add a column with the information, the taxonomical level at where it got it, and the literature source for that information.



```{r}
biom %>% 
  unite(Genus, Species,col = "species", sep = " ", remove = F) %>% 
  select(-Species) -> biom

biom %>% distinct (AdultMinerology)

#Aragonite -calcite and Calcite - aragonite


```

Let's see how many of the families are represented there

```{r}

summary(all.fams$family %in% biom$Family)

biom %>% 
  filter(Family %in% all.fams$Family) %>% 
  group_by(Family) %>% 
  summarise(n.sp  = n_distinct(Genus),
            n.ELS = n_distinct(EarlyLifeMinerology),
            n.ALS = n_distinct(AdultMinerology)) %>% 
  filter(n.ELS > 1 | n.ALS > 1) %>%  pull(Family) -> fams.with.diff.miner


```

We don't have information from 227 families, And from 15 families we have more than one way of mineralizing either larvae or adult. The approach we have followed is - if there is an annotation for its species, we take that; if there is a genus and if not, family.

```{r}
biom %>% 
  select(family = Family, genus = Genus, species, EarlyLifeMinerology, AdultMinerology, everything()) -> biom

sequential_join(df1 = all.hashes.found,
                 df2 = biom,
                 var.to.join.1 = "species" ,
                 var.to.join.2 = "genus" ,
                 var.to.join.3 = "family" ,
                 confidence = reliability.minerology,
                EarlyLifeMinerology, AdultMinerology) -> database.with.minerology

# For all families, add the order and class so we can get the in-between taxonomy


lineages=list(NA)
for (i in 1:length(unique( all.fams$family))){
        tempx=classification(unique( all.fams$family)[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

unique( all.fams$family)[-which(is.na(lineages))] -> fam.to.update

lineages[-which(is.na(lineages))] -> lineages



bind_rows(lineages, .id = "family") %>% 
  group_by(family) %>% 
  nest() %>% 
  mutate(family = fam.to.update) %>% 
  mutate(order = map(data, ~ .x %>%
                        #filter(rank == "order") %>% pull(name)))
                        filter(rank != "no rank") %>%
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>% 
  unnest(order) %>% 
  select(family, order, class) %>% 
  left_join(all.fams, .) -> all.fams

unique( all.fams$family)

# Now join the database with the taxonomy


database.with.minerology %>% 
  left_join(all.fams, by = "family") %>% 
  select(Hash, kingdom, phylum, class, order, family, genus, species, contains("tot."), contains("Minerology")) -> database.with.minerology


  
```
### Fill a few gaps in the taxonomy

Genus Dictyocha is a Stramenopiles frrom the phyllum Dictyochophyceae

```{r}
database.with.minerology %>% 
  filter(is.na(phylum), !is.na(genus)) %>% distinct(genus) %>% pull -> genus.to.lineage

genus.to.lineage

lineages=list(NA)
for (i in 1:length(genus.to.lineage)){
        tempx=classification(genus.to.lineage[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}


 bind_rows(lineages, .id = "genus") %>% # NOW CHANGE STRAMENOPILES TO PHYLUM
   mutate (rank = case_when(name == "Stramenopiles" ~ "kingdom",
                            TRUE ~ rank)) %>% 
   group_by(genus) %>%
   nest %>%
   mutate(genus = genus.to.lineage) %>%
   mutate(data2 = map(data, ~.x %>%
                        select(-id) %>%
                        filter(rank != "no rank") %>% 
                        spread(key = rank, value = name) ) ) %>%  
   unnest (data2) %>% 
   mutate(phylum = case_when(is.na(phylum) ~ class,
                             TRUE ~ phylum),
          phylum = case_when(is.na(phylum) ~ order,
                             TRUE ~ phylum)) %>% 
   select(genus, order, class, phylum, kingdom) -> fill.the.gaps
```

Now merge the new info

```{r}

database.with.minerology %>%  select(-kingdom, -phylum, -order, -class) %>% inner_join(fill.the.gaps, by = "genus") -> step1

database.with.minerology %>% anti_join(fill.the.gaps, by = "genus") %>% bind_rows(step1) -> database.with.minerology
```

# Step 2: add the info from trophic level

Trying to remember where did I get this file from. I think it comes from the MBON project. Check with Ryan. It comes from Collin

```{r}


trophic <- read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Data/TrophicData/subnetxtrophic_ordered_20181112_20181130.csv")

```

Check how many families are present in the dataset

```{r}
trophic %>%
  select(family = Rank5,
         
         order = Rank4,
         
         class = Rank3, 
         
         phylum = Rank2,
         
         trophic_level_fv, group_fv) -> trophic


trophic %>% 
  filter(is.na(Rank5)) %>%  distinct(Rank4) %>% pull -> NAs.family

lineages=list(NA)
for (i in 1:length(NAs.family)){
        tempx=classification(NAs.family[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

NAs.family[-which(is.na(lineages))] -> NAs.family

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "Rank4") %>% 
  group_by(Rank4) %>% 
  nest() %>% 
  mutate(Rank4 =  NAs.family) %>% # Make it an order
  mutate(last = map_chr (data, ~.x %>% last() %>% pull(name)))


```

No info for 256 families

```{r}
summary(all.fams$phylum %in% trophic$Rank4) # 4 by class

summary(all.fams$phylum %in% trophic$Rank3) # 51 by class

summary(all.fams$phylum %in% trophic$Rank2) # 299 by phyllum
```


Let's do a weird_join as before
```{r}

sequential_join(database.with.minerology,
                trophic,
                var.to.join.1 = "family",
                var.to.join.2 = "class",
                var.to.join.3 = "phylum",
                confidence = reliability.trophic,
                trophic_level_fv, group_fv) -> database_step.2
 
database_step.2 %>% 
  filter(reliability.trophic == "class")

```

# Step 3: Benthic or planktonic

OK, now the last step is to look for sources of life style, whether benthic or planktonic.

```{r load dino}

# First, a database from Gomez 2012

dinoflag <- read_csv("dinoflag_annotations.csv")
dinoflag %>% 
  select(family = Family, everything()) -> dinoflag
```

```{r load tara_oceans}
# Second, the dataset from the tara_oceans dataset - if it was on the first couple of meters, can we 
# assume it is a planktonic species?

tara_oceans <- read_delim("Tara_oceans.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

tara_oceans %>% 
  select(lineage, taxogroup, chloroplast, trophism) %>% 
  distinct() -> tara_oceans
tara_oceans %>% 
  filter(str_detect(lineage, pattern ="Amoebozoa" ))

 database.step.2 %>% distinct(kingdom) %>% arrange(kingdom) %>% filter(kingdom != "Metazoa") %>% pull() -> kingdoms.to.find.out

# Check how many species from our dataset are present in the Tara Oceans

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$species), collapse = "|"))) # Five spp

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$genus), collapse = "|"))) %>% 
  filter(str_detect(lineage, pattern = "Annelida"))# 451 genus

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$family), collapse = "|"))) # 289 families
```


```{r}
database_step.2 %>% 
  select(-n) -> database_step.2
sequential_join(database_step.2,
                dinoflag,
                var.to.join.1 = "species",
                var.to.join.2 = "genus",
                var.to.join.3 = "family",
                confidence = reliability.benthic,
                benthic, oceanic) -> database_step.3 


  
```

How many did get assigned?

```{r}
database_step.3 %>% 
  group_by(is.na(benthic)) %>% 
  summarise(nhash = n(),
            tot = sum(tot.reads))
```

58 Hashes, 362 thousand reads

```{r load massana2015}
Massana2015 <- read_csv("Massana2015.csv")

Massana2015 %>% 
  select(species = `Closest match to described species`, everything()) %>% 
  separate(species, into = "genus", sep = " ", remove = F) -> Massana2015

summary(med.step$genus %in% Massana2015$genus)

Massana2015 %>%
  group_by(genus) %>% 
  nest(species,Source) %>% 
  mutate(Ben = map_dbl (data, ~ sum(str_count(.x$Source, "Benthos"))),
         Plk = map_dbl (data, ~ sum(str_count(.x$Source, "plankton"))),
         Source = case_when(Ben == 0  ~ "PLK",
                            Plk == 0  ~ "BEN",
                            TRUE      ~ "Both")) %>% 
  select(genus, Source) -> Massana2015
```

```{r join Massana}
sequential_join(df1 = database_step.3,df2 = Massana2015, var.to.join.1 = "genus", var.to.join.2 = "",var.to.join.3 = "",confidence = confidence.source, Source) -> database_step.4


```


Let's see what is left to assign

```{r}
database_step.3 %>% 
  filter(is.na(benthic)) %>% 
  select(benthic,kingdom,phylum, class, order, family, genus, species ) %>% 
  distinct() %>% 
  arrange(kingdom)
  
```

Look for protist info from Forster2016. From there, turns out there is a database for planktonic protist - so let's look at it
```{r PFR2 }
PFR2_Metadata <- read_delim("PFR2_Metadata.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

PFR2_Metadata %>% 
  select(Genus, Species) %>% 
  unite(Genus, Species, col = "Species", sep = " ", remove = F) %>% 
  distinct(Species, .keep_all = T) -> PFR2_Metadata

lineages=list(NA)
for (i in 1:length(PFR2_Metadata$Genus)){
        tempx=classification(PFR2_Metadata$Genus[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

PFR2_Metadata$Genus[-which(is.na(lineages))] -> good.names

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "genus") %>% 
  group_by(genus) %>% 
  nest() %>% 
  mutate(genus = good.names) %>% 
  mutate(data2 = map (data, ~ .x %>% 
                        filter(rank %in% c("family", "order", "phylum")) %>% 
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>% 
  
  unnest(data2) -> test

# do they share any taxa?

summary(med.step$species %in% PFR2_Metadata$Species) # No species shared

unique(med.step$genus)[(unique(med.step$genus) %in% PFR2_Metadata$Genus)] # no genus shared -

unique (test$family) %in% unique(database_step.4$family)

# No Shared taxa here

```

None of them matched our samples
So, this seems like really complicated. Let's change the course of action. Print out the unique taxa/metadata as a csv file and fill the data with references.

```{r}
med.step %>%
  select(-Hash, -starts_with("tot.")) %>%
  distinct %>%
  select(kingdom, phylum, family, genus, species, benthic, benthic.reliability, everything()) %>%
  arrange(desc(kingdom )) %>% filter(is.na(family))
#   write_csv("taxa.for.life.style.csv")
```

Done some editing - see if there is a big chunk I can do from the Massana work

```{r}
med.step.summ <- read_csv("taxa.for.life.style.csv")

med.step.summ %>% 
  filter (str_detect(phylum, "Florideoph"))

med.step.summ %>% 
  filter(is.na(benthic)) %>% 
  left_join(Massana2015, by = "genus") %>% 
  filter(!is.na(Source)) %>% 
  mutate(benthic = Source,
         benthic.reliability = "genus",
         source.benthic = "Massana2015") %>% 
  select(  - Source ) -> to.bind
  
med.step.summ %>% #filter(is.na(family))
  filter(!is.na(benthic)) %>% 
  bind_rows(to.bind) %>% 
  bind_rows(med.step.summ %>% 
              filter(is.na(benthic)) %>%
              anti_join(Massana2015, by = "genus")) %>%
  write_csv("taxa.for.life.style2.csv")

```
Emily Jacobs-Palmer found another source of benthic / pelagic diatoms

```{r loading Kasim2006}

Kasim2006 <- read_csv("Kasim_Mukai2006")



# Separate No. and Name of species
Kasim2006 %>% 
  separate(`No. Name of species`, into = c("No", "Species"), sep = " ", extra = "merge") %>% 
  separate(Species, into = c("genus","species"), sep = " ", remove = F) %>%
  unite(genus, species, col = "species", sep = " ", remove = F) %>% 
  mutate(Source = case_when(Group == "B" ~ "BEN",
                            Group == "P" ~ "PLK")) -> Kasim2006

#Do species from the same genus share lifestyle?

Kasim2006 %>% 
  group_by(genus) %>% 
  summarise(n = n_distinct(Source)) %>% 
  filter (n>1) # No,so I can simplify the table


Kasim2006.gen <- Kasim2006 %>% 
  group_by(genus) %>% 
  summarise(Source = first(Source))

# Done - now get the lineages of those genus

lineages=list(NA)

for (i in 1:length(Kasim2006.gen$genus)){
        tempx=classification(Kasim2006.gen$genus[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

Kasim2006.gen$genus[-which(is.na(lineages))] -> genus.to.update

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "genus") %>% 
  group_by(genus) %>% 
  nest() %>% 
  mutate(genus = genus.to.update) %>% 
  mutate(data2 = map (data, ~ .x %>% 
                        filter(rank %in% c("family", "order", "phylum")) %>% 
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>% 
  
  unnest(data2) -> test



Kasim2006.gen %>% 
  inner_join(test %>% select(-data), by = "genus") -> Kasim2006.gen

Kasim2006.gen %>% 
  select(Source.Kasim = Source, everything()) -> Kasim2006.gen

```


Now do the sequential join with Kasim

```{r join Kasim}
database_step.4 %>% 
  select(-n) -> database_step.4


sequential_join(df1 = database_step.4,
                df2 = Kasim2006.gen,
                var.to.join.1 = "genus",
                var.to.join.2 = "family",
                var.to.join.3 = "order",
                confidence = "source.Kasim",
                Source.Kasim) -> database_step.5

# Let's load our personal database - keeping only the benthic info

med.step.2 <- read_csv("taxa.for.life.style2.csv")

summary(med.step.2$family %in% database_step.5$family)

summary(is.na(med.step.2$family))

med.step.2 %>% 
  filter(is.na(family)) %>% 
  distinct(genus) %>% 
  semi_join(database_step.5, by = "genus")

# So all the families are there, and for those that are not, we have the genus - we can join it with database5
```

```{r join csv with data5}

# 1 remove n

database_step.5 %>% 
  select(-n) -> database_step.5

med.step.2 %>% 
  select(Source.csv = benthic, everything()) %>% 
  sequential_join(df1 = database_step.5, 
                  df2 = ., 
                  var.to.join.1 = "species", 
                  var.to.join.2 = "genus", 
                  var.to.join.3 = "family", 
                  confidence = "Summary.benthic",
                  Source.csv,source.benthic ) %>% distinct() -> database_step.5 
 
  


```
Now that I have all the benthic information in a table, let's deal withthe disagreemnts

```{r differences in benthic assignment}

# Step1
# for which phylum do we have an answer

database_step.5 %>%
  filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  distinct(phylum) # $43 phyla

# Step2
# disagreements

database_step.5 %>%
  #filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  unite(benthic, Source, Source.Kasim, Source.csv, sep = "|", col = all.benthos) %>% 
  mutate(agreed.benthos = case_when(str_detect(all.benthos, "PLK") ~ "PLK",
                                    str_detect(all.benthos, "BEN") ~ "BEN",
                                    str_detect(all.benthos, "Both")~ "Both" ,
                                    TRUE                           ~ "None")) %>% 
  select(Hash, kingdom, phylum, class, order, family, genus, species, contains("minerology"), contains("_fv"), benthos = agreed.benthos
         ) -> to.export

  

```

Collapse some values of the group_fv variable

```{r fixing group_fv}
write_csv(to.export, path = "hash.annotated.csv")


read_csv("hash.annotated.csv") %>% 
  
  
  unite (family, genus, species, col ="taxa", sep = "|", remove = F) %>% 
  
  mutate(group_fv = case_when(group_fv == "daitom" ~ "diatom",
                              group_fv == "copepods" ~ "copepod",
                              class == "Clitellata"  ~ "earthworm",
                              phylum == "Nematoda"   ~ "worm",
                              str_detect(taxa, "Harpacticoida") ~"copepod",
                              str_detect(taxa, "Polychaeta") ~"polychaete",
                              group_fv == "amoeboid" ~ "amoebozoa",
                              str_detect(taxa, "Parvamoeba") ~ "amoebozoa",
                              str_detect (group_fv, "echino") ~ "echinoderm",
                              str_detect (group_fv, "flagellate") ~ "flagellate",
                              str_detect(taxa, "Dictyocha|Pseudochatt") ~ "flagellate",
                              str_detect (group_fv, "hydrozoa") ~ "hydrozoan",
                              str_detect (taxa, "Campanulariidae") ~ "hydrozoan",
                              !is.na(group_fv) ~ group_fv,
                              str_detect(taxa, "Dinophysiaceae|Prorocentra|Gonyaulacaceae|Gymnodiniaceae|Kareniaceae|Thoracosphaeraceae") ~ "dinoflagellates",
                              str_detect(taxa, "Skeletonemataceae|Bacillariaceae|Stephanodiscus|Naviculace|Chaetoceros") ~ "diatom",
                              str_detect(taxa, "Cymbomonas|Haramonas|Trebouxia|Pycnoc") ~ "algae, green",
                              str_detect(taxa, "Rhodomelace") ~ "algae, red",
                              kingdom == "Rhodophyta" ~ "algae, red",
                              str_detect(taxa, "Aspergillaceae|Phytop|Haptoglo|Pythia|Pyramimonas|Saproleg|Apodachlya") ~ "fungi",
                              kingdom == "Stramenopiles" ~ "fungi" )) %>% 
  select(-taxa) -> to.export 
  
  write_csv(to.export, path = "hash.annotated.csv")



```
```{r aragonite calcite}
 to_export <- read_csv("hash.annotated.csv")

to_export %>% 
 # slice(1:10) %>% 
  mutate_at(vars(contains("Minerology")), function (x) case_when(x == "Calcite, Aragonite" ~ "Aragonite, Calcite", 
                                                                 x == "Caclite, Aragonite" ~ "Aragonite, Calcite",
                                                                 TRUE                      ~ x)) %>% 
 write_csv("hash.annotated.csv")
```
Fixing some mistakes

```{r}
to_export <- read_csv("hash.annotated.csv")

# Fix the polychaete Sccocirrus
to_export [which (to_export$species == "Saccocirrus sp. CMC01"), c(9:11, 13)] <- list("Non-CaCO3 mineral", "Non-CaCO3 mineral", "family", "polychaete")
# fix some undtermined
to_export[which (to_export$Hash == "fc9fed56b0203902b5eff4eca28b358d4419813a"),12:14] <- list("fungi", "fungi", "None"  )

to_export[which (to_export$Hash == "0124f9c1e949536ce587cd038df6e6f635a8a7d8"),12:14] <- list("3", "copepod", "PLK"  )    
to_export[which (to_export$Hash %in%c("459872f12aa1cf78e8eafa97db881ba6ef6af564",
                                      "b074016d069e564387f67ad18a6e5bc181726411",
                                      "60e1a14acfe5ff45c2b2dbf5b29bb067632a3bf3",
                                      "d6a7fde26da9282448c961966b6d50ab554e76a5",
                                      "2f8238e8b72201b38768d46b32b1266ddc892f0c") ),12:14] <- list("3, worm", "polychaete", "BEN"  )

to_export[which (to_export$Hash %in%c(  "3b86a8af70ee3ae734e65aed41f36d6169e3f60a",
                                        "a14d493abcb3b525eb449d241bd2564ca50eba6f",
                                        "448756e5f9def73770c548b6b772d9d38d4f5bc2") ),12:14] <- list("3, insect", "undetermined", "None"  )                         
# Fill in some taxonomy

  to_export[which(to_export$species == "Harpacticoida sp. DNZ085"), 2:5]  <- list("Metazoa", "Arthropoda", "Hexanauplia", "Harpacticoida")               
    
  to_export[which(to_export$species == "Polychaeta sp. BOLD:ACQ0474"), 2:4] <- list("Metazoa","Annelida","Polychaeta")
  to_export[which(to_export$species == "Polychaeta sp. NB-Po356"), 2:4] <- list("Metazoa","Annelida","Polychaeta")
   to_export[which(to_export$species == "Psocoptera sp. BOLD:ACB0984"), 2:4] <- list("Metazoa","Arthropoda","Insecta")      

# Some copepods too   
   to_export[which(to_export$species %in%c("Centropages abdominalis","Nitokra spinipes",
                                        "Stenhelia pubescens") ), 13] <- "copepod"

 to_export[which(to_export$class == "Polychaeta"), "group_fv"] <- "polychaete"
 
 to_export %>% 
    write_csv(here("Input", "hash.annotated.csv"))

```


```{r write.out}

to.export %>% 
  filter (is.na(group_fv)) %>% 
  select (-Hash) %>% 
  distinct() %>% 
  arrange(kingdom, phylum)

```

```{r summary stats}

database_step.5 %>%
  #filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  unite(benthic, Source, Source.Kasim, Source.csv, sep = "|", col = all.benthos) %>% 
  mutate(agreed.benthos = case_when(str_detect(all.benthos, "PLK") ~ "PLK",
                                    str_detect(all.benthos, "BEN") ~ "BEN",
                                    str_detect(all.benthos, "Both")~ "Both" ,
                                    TRUE                           ~ "None")) %>% 
  group_by(agreed.benthos) %>% 
  summarise(nReads = sum(tot.reads),
            nHash = n()            )

```

